
Loading Packages
```{r results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(ggrepel)
library(dplyr)
```

Importing data
```{r}
github_Link <-"https://raw.githubusercontent.com/EvanLih/PUBPOL-543-BEES-Project/master/Final_Data.csv"
finalData <- read.csv(github_Link)
finalData = finalData[-c(161:999),] # Cleaning
```

## <font color="red">Univariate Analysis</font>
### Plotting Electricity Access *(Density Plot)*

**A summary of the global distribution of electricity access.**

Prepare the data frame:
```{r}
finalData$accessElectricity=as.numeric((finalData$accessElectricity))
df1=data.frame(finalData$Country,finalData$accessElectricity,finalData$RenewConsumption) # Creating the data frame for Electricity Access & Renewable Consumption
```
Plot:
```{r}
base1= ggplot(df1,aes(x = finalData$accessElectricity))  
base1= base1 + stat_density(geom="line",color="Red") +  labs(x = "Access to Electricity (% of population))", y = "Density",
       caption = "Source : The World Bank (2015)",
       title = "What is The Global Distribution of Electricity Access?")   + theme(plot.caption = element_text(hjust= -0.2))+ theme_stata()
base1 
```

**Main Findings:**

- The density plot a longer tail to the left, which means that the majority of countries have high accessibility to electricity.
- There are lots of countries have 100% capacity to electricity.

### Plotting Renewable Consumption *(Density Plot)*

**A summary of the global distribution of renewable energy consumption.**

Plot:
```{r}
base2= ggplot(df1,aes(x = finalData$RenewConsumption))  
base2= base2 + stat_density(geom="line",color="Red") +  labs(x = "Renewable Consumption (% of total energy consumption)", y = "Density",
       caption = "Source : The World Bank (2015)",
       title = "What is The Global Distribution of Renewable Consumption?")    + theme(plot.caption = element_text(hjust= -0.2))+ theme_stata()
base2 
```

**Main Finding:**

- The highest point of curve is located at around 10%, which means a great number of countries have a low renewable consumption level.

### Plotting Land Use *(Grouped Bar Chart)*

**A summary of the continental average utilization rate for different land types.**

Land types contain agricultural land *(lands being used for agriculture)*, arable land *(lands suitable for agriculture)* and forest lands. By calculating the differece between agricultural and arable land, we aim to see ***how much of the land that wasn't meant for agriculture has been turned into agriculture*** and thus learning the average land development sustainability for each continent. 

Prepare the data frame:
```{r} 
ContinentLand = finalData %>% select(c("Continent", "Country","Agricultural","Arable.Land","Forest")) 
Landdf = as.data.frame(ContinentLand)
Landdf$`Arable.Land`=as.numeric((Landdf$`Arable.Land`)) 
Landdf$`Arable.Land`=round(Landdf$`Arable.Land`,digit=2) 
```
Calcultae the difference between agricultural land and arable land, and create a new variable for the difference:
```{r}
differenceLand=c(Landdf$Agricultural-Landdf$`Arable.Land`)
Landdf$diff = differenceLand 
```
Calculate continental average for each variable and update the data frame:
```{r}
AgriMean = aggregate(data=Landdf, Agricultural~Continent,mean)
AraMean = aggregate(data=Landdf, Arable.Land~Continent,mean)
ForMean = aggregate(data=Landdf, Forest~Continent,mean)
diffMean = aggregate(data=Landdf, diff~Continent,mean)
LandMean = data.frame(AgriMean,AraMean,ForMean,diffMean)
LandMean = LandMean[,-c(3,5,7)] 
```
```{r}
library(reshape2)
LMm = melt(LandMean) 
```
Plot:
```{r}
base3 = ggplot(LMm,aes(x = Continent,y = value)) + 
  labs(x = "Continent", 
       y = "Average Land Utilization Rate",
       caption = "Source : The World Bank (2015)",
       title = "How Do Land Sustainable Development Levels Compare by Continent?") +
  theme(plot.caption = element_text(hjust= -0.2)) +
  geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +
  scale_fill_brewer(labels = c("Agricultural Land", "Arable Land","Forest Land","Difference between Agricultural & Arable Land"),
    palette="Pastel1") + 
  theme_stata() + 
  guides(fill = guide_legend(title = "Land Type", title.position = "top",nrow = 2))

base3
```

**Main Findings:**

- All continents over exploited the arable land resource.
- Europe is doing better at sustainable land use.
- Africa and South America are doing worse.

### Plotting Income *(Bar Chart)*
 
**A summary of the number of countries with different income levels in each continent.**

Each country is categoried as low, lower middle, upper middle or high income country, and we count the number of countries categoried as different income groups in each continent. 

Prepare the data frame:
```{r}
ContinentIncomeGroup = finalData %>% select(c("Continent", "Country","Income.Group"))
df2 = as.data.frame(ContinentIncomeGroup) # Creating the data frame
df2$Income.Group = factor(df2$Income.Group,levels = c("Low income","Lower middle income","Upper middle income","High income")) # Reordering income categories
```
Plot:
```{r}
base4 = ggplot(df2, aes(x = Income.Group)) +
  geom_bar(aes(y = (..count..)/sum(..count..))) +
  labs(x = "Income Group", 
       y = "Proportion",
       caption = "Source : The World Bank (2015)",
       title = "How Do Income Levels Compare by Continent?") +
  scale_y_continuous(labels = scales::percent, name = "Proportion") +
  theme(plot.caption = element_text(hjust= -0.2)) +
  facet_wrap(~ Continent) + coord_flip() + theme_stata() +
  theme(axis.text.y = element_text(size = 6, angle = 45)) 

base4
```

**Main Findings:**

- Europe has a significant proportion of high-income countries.
- Africa is dominated by low- & middle- income countries.
- Income level of Asian countries varies a lot.


## <font color="red">Bivariate Analysis</font>
### Ploting SDI and Land Use 

**A comparison of SDI and land development sustainability (the difference between agricultural land and arable land).**

#### by Country *(Hexagonal Heatmap)*
Prepare the data frame:
```{r}
SDIByCountry=data.frame(finalData$Country,finalData$`Sustainable.Development.Index`)
names(SDIByCountry)=c('Country','SDIScore')
LandByCountry=data.frame(finalData$Country,differenceLand)
names(LandByCountry)=c('Country','LandUse')
```
```{r}
num_num=merge(SDIByCountry,LandByCountry) 
num_num$Country=as.character(num_num$Country)
```
Plot:
```{r}
SDILANDCountry = ggplot(num_num, aes(y = SDIScore, x = LandUse)) + 
  labs(x = "Difference between Agricultural Land & Arable Land", 
       y = "Sustainability Development Index",
       caption = "Source : Sustainable Development Project and The World Bank (2015)",
       title = "Is Land Sustainable Development Level Related to SDI?") +
  theme(plot.caption = element_text(hjust= -0.2)) +
  theme_stata()

# highlighting overlaping of points
SDILANDCountry = SDILANDCountry +  
  geom_hex(bins = 10) +
  scale_fill_distiller(palette ="Greys",direction=1) +
   guides(fill = guide_legend(title = "Count", title.position = "top",size=1)) +
  theme(legend.position="right")
SDILANDCountry
```

**Main finding:**

- The darker color in the upper left corner indicates that a great number of countries are in that area. From this we can conclude a positive relationship between SDI and land development sustainabiity level. That is, the smaller difference between agricultural & arable land, the higher level of land development sustainability, and the higher level of SDI. 

#### by Continent *(Scatterplot)*
Calculate continental average of data, and prepare the data frame:
```{r}
# Comparing average continental SDI score and land susatinability level
SDIByContinent=aggregate(data=finalData, `Sustainable.Development.Index`~Continent, mean)
LandUseByContinent=aggregate(data=finalData, differenceLand~Continent, mean)
```
```{r}
num_num1=merge(SDIByContinent,LandUseByContinent) 
num_num1$Continent=as.character(num_num1$Continent)
```
Plot:
```{r}
SDILANDContinent = 
  ggplot(num_num1, aes(y = `Sustainable.Development.Index`, x = differenceLand)) + 
  geom_point() + 
  geom_text_repel(aes(label=Continent),size=2.5) +
  labs(x = "Difference between Agricultural Land & Arable Land", 
       y = "Average Sustainability Development Index Rating",
       caption = "Source : Sustainable Development Project and The World Bank (2015)",
       title = "Is Continental Land Sustainable Development Level Related to SDI?")+
  theme(plot.caption = element_text(hjust= -0.2))+
  theme_stata()
SDILANDContinent
```

**Main findings:**

- Europe has high land development sustainability level and low SDI.
- South America has low land development sustainability level and high SDI.
- Africa has low land development sustainability level and relatively low SDI.
- The continental comparison result is not consistent with the result in country level. We guess it's partially due to our limitations in data processing. For example, when comparing continental average data, it's better for us to calculate weighted mean instead of absolute mean, so that we can take into account more factors such as country size and GDP.
